---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# clggm 

`clggm` contatins methods to estimate the coefficients of the stochastic 
process (Ornstein–Uhlenbeck)
$dX_t = B(X_t - a)dt + DdW_t$
from observations of the invariant distribution.

## Installation 

```{r, eval = FALSE}
devtools::install_github("gherardovarando/clggm")
```

### Dependencies 

- `MASS`
- `stats` 
- `testthat` (only in dev)


## Usage

```{r}
library(clggm)

### define coefficient matrices
B <- matrix(nrow = 4, c(-4, 1,   0,  0, 
                         0, -3,  1,  0,
                         0,  0, -2,  1,
                         0,  0,  0, -1), byrow = TRUE)
D <- diag(c(1,2,1,2))
C <- D %*% t(D)

### solve continuous Lyapunov equation 
### to obtain covariance of invariant 
### distribution
Sigma <- clyap(B, C) 

### obtain observations from the invariant distribution
sample <- MASS::mvrnorm(n = 10000, mu = rep(0,4),Sigma =  Sigma)
``` 


### Estimate $B$ knowing $C$

```{r}
B0 <- - 0.5 * C %*% solve(cov(sample))
### penalized maximum-likelihood solved with proximal gradient
results <- proxgradllB(cov(sample), B= B0, C = C, eps = 1e-12, 
                       maxIter = 1000, lambda = 0.05, job = 1)
results 

### penalized least square 0.5*||S(B) - Sigma||_2^2 + lambda * ||B||_1,off
results2 <- proxgradlsB(cov(sample), B= B0, C = C, eps = 1e-12, 
                       maxIter = 1000, lambda = 0.01, job = 0)
results2$B
```

#### Solutions along a regularization path

```{r}
results <- llBpath(Sigma, eps = 1e-12, maxIter = 1000, job = 11)
t(sapply(results, function(res) c(lambda = res$lambda, 
                                npar = sum(res$B!=0),
                                fp = sum(res$B!=0 & B==0),
                                tp = sum(res$B!=0 & B!=0) ,
                                fn = sum(res$B==0 & B!=0),
                                tn = sum(res$B==0 & B==0),
                                errs = sum(res$B!=0 & B==0) + 
                                  sum(res$B==0 & B!=0))))
```


### Estimate $C$ knowing $B$

```{r}
results <- graddsllc(cov(sample), B, C = diag(4), C0 = diag(4), 
                      eps = 1e-12, maxIter = 1000, lambda = 0.0005)
cbind(C = diag(C), Cest = diag(results$C))
```

### Estimate $B$ and $C$

```{r}
B0 <- - 0.5 * diag(4) %*% solve(cov(sample))
results <- pnllbc(cov(sample), B0, C = diag(4), C0 = diag(4), 
                      eps = 1e-14, maxIter = 1000, intitr = 1,
                      lambda = 0.2, lambdac = 0.005, job = 0)
## B estimated
results$B

## C estimated
cbind(C = diag(C), Cest = diag(results$C))
```

## Related code

- Some code is from the `lyapunov` package (https://github.com/gragusa/lyapunov).

- Using FORTRAN code of Algorithm 705 from the Collected Algorithms from ACM, TRANSACTIONS ON MATHEMATICAL SOFTWARE, VOL. 18, NO. 2, PP. 232-238.75.